### 2024-01-22 code/exploratory/data_exploration_v1.Rmd
region identification, figure 2 adjustments

---
title: "SCBI 2023 Census Exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "hide",message = FALSE, warning = FALSE)
```

```{r init}
library(tidyverse)
library(data.table)
library(sf)
library(RColorBrewer)

curStem <- fread("C:/Work/Smithsonian/Repos/2023census/processed_data/scbi.stem4.csv")

prevStem <- fread(paste0("https://raw.githubusercontent.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/master/tree_main_census/data/census-csv-files/scbi.stem3.csv"))

load("C:/Work/Smithsonian/Repos/Morreale_BiomassChange/data/census/scbi.stem2.rdata")

load("C:/Work/Smithsonian/Repos/Morreale_BiomassChange/data/census/scbi.stem1.rdata")

quadrats <- st_read( "C:/Work/Smithsonian/Repos/2023census/doc/maps/20m_grid/20m_grid.shp")

```

### Recruitment 2018-2023 ###

#### Trees ####

```{r recruits I}

recruits <- curStem %>% 
  filter(curStem$table == "recruits") %>% 
  select(tag,StemTag,quadrat,sp)

r_stems <- recruits %>% 
  group_by(sp) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
r_trees <- recruits %>% 
  group_by(tag) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(sp) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

ggplot(r_trees,aes(x = reorder(sp,-n),y = n, fill = -n)) +
  geom_bar(stat =  "identity",col = "black") +
  scale_fill_gradient(low = "grey20",high = "grey80") +
  theme_classic() +
  ylab("# of Recruits (Trees)") +
  xlab(" ") +
  theme(legend.position = "none")
```

#### Stems ####

```{r recruits II}
ggplot(r_stems,aes(x = reorder(sp,-n),y = n, fill = -n)) +
  geom_bar(stat =  "identity",col = "black") +
  scale_fill_gradient(low = "grey20",high = "grey80") +
  theme_classic() +
  ylab("# of Recruits (Stems)") +
  xlab(" ") +
  theme(legend.position = "none")
```

### Mortality 2018-2023 ###

#### Stems ####

```{r mortality}
ids_prev_living <- prevStem %>% 
  filter(status == "A") %>% 
  mutate(taglink = paste0(tag,StemTag))
m_stems <- curStem %>% 
  mutate(taglink = paste0(tag,StemTag)) %>% 
  filter(taglink %in% ids_prev_living$taglink & status_current != "LI") %>% 
  group_by(sp) %>% 
  summarize(n = n())

ggplot(m_stems,aes(x = reorder(sp,-n),y = n, fill = -n)) +
  geom_bar(stat =  "identity",col = "black") +
  scale_fill_gradient2(low = "grey40",high = "grey80") +
  theme_classic() +
  ylab("Mortality (# of Stems)") +
  xlab(" ") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))

```

### Compositional Shifts ###

#### Relative Abundance (by stem count) of top 15 species in 2018 census ####
```{r }
complete_quads <- unique(curStem$quadrat) %>% na.omit()

c2018_ra <- prevStem %>% #calculate relative abundance for 2018 census in quadrats that have been completed so far in 2023
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  group_by(sp) %>% 
  summarize(StemCount = n()) %>% 
  mutate(RelAbund = StemCount/sum(StemCount), Census = "2018") %>% 
  arrange(desc(RelAbund))

c2013_ra <- scbi.stem2 %>%  
  mutate(quadrat = as.numeric(quadrat)) %>% 
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  group_by(sp) %>% 
  summarize(StemCount = n()) %>% 
  mutate(RelAbund = StemCount/sum(StemCount), Census = "2013") %>% 
  arrange(desc(RelAbund))

c2008_ra <- scbi.stem1 %>%  
  mutate(quadrat = as.numeric(quadrat)) %>% 
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  group_by(sp) %>% 
  summarize(StemCount = n()) %>% 
  mutate(RelAbund = StemCount/sum(StemCount), Census = "2008") %>% 
  arrange(desc(RelAbund))

allCens_ra <- curStem %>% 
  filter(status_current == "LI") %>% 
  group_by(sp) %>% 
  summarize(StemCount = n()) %>% 
  mutate(RelAbund = StemCount/sum(StemCount), Census = "2023") %>% 
  bind_rows(c2018_ra, c2013_ra,c2008_ra) %>% 
  #bind_rows(c2018_ra) %>% 
  mutate(Census = as.factor(Census))

topRanked_sp <- c2018_ra %>% 
  slice_head(n = 15)

plotdf <- allCens_ra %>% 
  filter(sp %in% topRanked_sp$sp)

ggplot(plotdf,aes(x = reorder(sp,-RelAbund),y = RelAbund, group = Census, fill = Census)) + 
  geom_bar(stat = "identity",position = "dodge", col = "black") +
  scale_fill_grey() +
  xlab("") +
  ylab("Relative Abundance") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45))
```

### Species Size Class Distributions of top 15 species in 2018 census ###

```{r fig.height= 30, fig.width = 6}
brks <- c(seq(0,120, by = 20),170)

c2018_sp_size_ab <- prevStem %>%  
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  mutate(dbh_classes = cut(as.numeric(dbh)/10,breaks = brks,right = F)) %>% 
  group_by(sp,dbh_classes) %>% 
  summarize(StemCount = n()) %>% 
  ungroup() %>% 
  complete(sp,nesting(dbh_classes),fill = list(StemCount = 0)) %>% 
  mutate(Census = "2018") %>% 
  arrange(desc(StemCount))

c2013_sp_size_ab <- scbi.stem2 %>%  
  mutate(quadrat = as.numeric(quadrat)) %>% 
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  mutate(dbh_classes = cut(as.numeric(dbh)/10,breaks = brks,right = F)) %>% 
  group_by(sp,dbh_classes) %>% 
  summarize(StemCount = n()) %>% 
  ungroup() %>% 
  complete(sp,nesting(dbh_classes),fill = list(StemCount = 0)) %>% 
  mutate(Census = "2013") %>% 
  arrange(desc(StemCount))

c2008_sp_size_ab <- scbi.stem1 %>%  
  mutate(quadrat = as.numeric(quadrat)) %>% 
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  mutate(dbh_classes = cut(as.numeric(dbh)/10,breaks = brks,right = F)) %>% 
  group_by(sp,dbh_classes) %>% 
  summarize(StemCount = n()) %>% 
  ungroup() %>% 
  complete(sp,nesting(dbh_classes),fill = list(StemCount = 0)) %>% 
  mutate(Census = "2008") %>% 
  arrange(desc(StemCount))

allCens_sp_size_ab <- curStem %>% 
  filter(status_current == "LI") %>% 
  mutate(dbh_classes = cut(as.numeric(dbh_current)/10,breaks = brks,right = F)) %>% 
  group_by(sp,dbh_classes) %>% 
  summarize(StemCount = n()) %>% 
  ungroup() %>% 
  complete(sp,nesting(dbh_classes),fill = list(StemCount = 0)) %>% 
  mutate(Census = "2023") %>% 
  bind_rows(c2018_sp_size_ab, c2013_sp_size_ab,c2008_sp_size_ab) %>% 
  #bind_rows(c2018_sp_size_ab) %>% 
  mutate(Census = as.factor(Census))

plotdf_size <- allCens_sp_size_ab %>% 
  filter(sp %in% topRanked_sp$sp)

ggplot(plotdf_size,aes(x = dbh_classes,y = StemCount, group = Census, fill = Census)) + 
  geom_bar(stat = "identity",position = "dodge", col = "black") +
 # geom_line(aes(col = Census)) +
  facet_wrap(~sp,scales = "free",ncol = 1) +
  scale_fill_grey() +
  xlab("") +
  ylab("Number of stems") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45,vjust = .8))

```

### Change in Aboveground Biomass ###

```{r}
library(allodb)

sp.table <- fread(paste0("https://raw.githubusercontent.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/master/species_lists/Tree%20ecology/SCBI_ForestGEO_sp_ecology.csv")) %>% 
  select(spcode,family,genus,species)
latlong <- c(-78.1454, 38.8935)
area_ha_measured <- (length(unique(curStem$quadrat)) * 400) / 10000

curLive <- curStem %>% 
  filter(status_current == "LI") %>% 
  select(quadrat,dbh = dbh_current,spcode = species) %>% 
  mutate(Census = "2023")
AllLive <- scbi.stem1 %>%  
  mutate(Census = "2008") %>% 
  bind_rows(scbi.stem2 %>% mutate(Census = "2013")) %>% 
  mutate(quadrat = as.numeric(quadrat)) %>% 
  filter(status == "A" & quadrat %in% complete_quads) %>% 
  select(quadrat,dbh, spcode = sp,Census) %>% 
  bind_rows(prevStem  %>% 
              filter(status == "A" & quadrat %in% complete_quads) %>% select(quadrat,dbh,spcode = sp) %>% mutate(Census = "2018")) %>% 
  mutate(dbh = as.numeric(dbh)) %>% 
  bind_rows(curLive)

AGB_stems <- AllLive %>% 
  left_join(sp.table) %>% 
  mutate(AGB = allodb::get_biomass(dbh = dbh/10, genus = genus, species = genus, coords = latlong))

census_AGB <- AGB_stems %>% 
  mutate(AGB_Mg = AGB / 1000) %>% 
  group_by(Census) %>% 
  summarize(AGB_ha = sum(AGB_Mg)/area_ha_measured)

species_AGB <- AGB_stems %>% 
#  filter(spcode %in% topRanked_sp$sp) %>% 
  mutate(plt_sp = if_else(spcode %in% c("litu","fram","qual","quru","quve"),
                          spcode,"Other")) %>% 
  mutate(AGB_Mg = AGB / 1000) %>% 
  group_by(Census,plt_sp) %>% 
  summarize(AGB_ha = sum(AGB_Mg)/area_ha_measured)

ggplot(census_AGB ,aes(x = as.numeric(Census), y = AGB_ha)) +
  geom_point(cex = 2, col = "blue") +
  geom_line(alpha = .2, lwd = 1.2, col = "blue") +
  xlab("Census") +
  ggtitle("SCBI Aboveground Biomass") +
  ylab(expression(AGB~(Mg~ha^-1))) +
  theme_classic() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20),
        title = element_text(size = 22))

ggplot(species_AGB ,aes(x = as.numeric(Census), y = AGB_ha,group = plt_sp, col = plt_sp)) +
  geom_point(cex = 2) +
  geom_line(alpha = .2, lwd = 1.2) +
  xlab("Census") +
  ggtitle("Selected Species AGB") +
  ylab(expression(AGB~(Mg~ha^-1))) +
  scale_y_continuous(breaks = seq(0,140,by = 20))  +
  scale_color_brewer(palette = "Dark2", name = "Species") +
  theme_bw() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20),
        title = element_text(size = 22))

```